\section{Context}

\begin{frame}
	\frametitle{Context}
	\pipelinesDashboard
	\note{"This is a tool I use for my research." Just briefly show the nature of the data.
		\begin{itemize}
			\item We have comprehensive data on pipelines in the US.
			\item We know which organizations have how many miles. (show top 5)
			\item We know how old their pipelines are.
			\item We know how many incidents occur.
			\item So it would be a great context to study strategy, and how it affects incident rates.
		\end{itemize}}
\end{frame}

\begin{frame}
	\frametitle{Cases}
	\incidentDashboard
	\note{"This is a tool I use for my research." Just briefly show the scale, because i found it hard to imagine, just from numbers.
		\begin{itemize}
			\item Show all organizations at once, add some jitter, alpha.
			\item Show one group, remove jitter, alpha.
		\end{itemize}}
\end{frame}

\begin{frame}
	\frametitle{Context}
	\begin{block}{High-profile bankruptcies related to pipeline incidents}
		\begin{itemize}
			\item Greka Energy/HVI Cat Canyon (1996, 2005, 2019)
			\note{Greka: History of oil spills, settlements relatively unexpensive, e.g., \$12 million fine caused bankruptcy in 2019, ironically, CEO once stated company motto as "Working for profits".\medbreak}
			\item Pacific Gas and Electric Company (2001, 2019)
			\note{PG\&E: After 2010 explosion with 8 deaths, paid \$300m in fines, \$400m in refunds, \$850m for upgrades, and \$500m in settlements.\medbreak}
			\item{EdgeMark (2019)}
			\note{EdgeMark: Energy Transfer pipeline explosion led to EdgeMark Bankruptcy. Ontario Teachers' Pension fund held minority stake. \bigbreak}
		\end{itemize}	
	\end{block}

	\hrulefill
	
	$\rightarrow$ There is some relevance for traditional performance also.
	\note{\hrulefill \smallbreak
		I don't explore this relevance for performance any further though.}
\end{frame}
