\section{Methods}

\begin{frame}
	\frametitle{Before I forget...}
	\begin{itemize}
		\item Fixed-effect
		\item Company-year
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Independent variable}
	Change of the organization over time - pipeline assets.\smallbreak
	$\rightarrow$ \textbf{Low} if the organization is not changing, or changing at constant pace.\smallbreak
	$\rightarrow$ \textbf{High} if organization deviates from previous course.
	
	\hrulefill
	
	$A_{t} = ((C_{t} - \bar{C})^2 + (C_{t-1} - \bar{C})^2 + (C_{t-2} - \bar{C})^2) / 2$\smallbreak
	where\smallbreak
	$C = M_{t} - M_{t-1}$\smallbreak
	$M$ is \textbf{M}iles of pipeline\smallbreak
	$A$ is strategic \textbf{A}djustments\smallbreak
\end{frame}

\begin{frame}
	\frametitle{Dependent variable}
	Incidents - three year average
	
	\hrulefill
	
	$\bar{I} = (I_{t} + I_{t+1} + I_{t+2})/3$\smallbreak
	where\smallbreak
	$I$ is number of significant incidents
\end{frame}

\begin{frame}
	\frametitle{Dependent variable}
	Fulfill any of these four criteria:
	\begin{itemize}
		\item "Fatality or injury requiring in-patient hospitalization
		\item \$50,000 or more in total costs, measured in 1984 dollars
		\item Highly volatile liquid releases of 5 barrels or more or other liquid releases of 50 barrels or more
		\item Liquid releases resulting in an unintentional fire or explosion
	\end{itemize}
	\incidentDefinition
\end{frame}

\begin{frame}
	\frametitle{Data source}
	Pipeline and incident data extracted from two different datasets from FERC (Federal Energy Regulatory Commission).
	\begin{itemize}
		\item Same dataset (incidents): \citet{Park2019}
		\note{FERC frequently used in econometrics/economics papers, e.g., BEPP.}
	\end{itemize}
	M\&A, ownership data obtained from LexisNexis.
\end{frame}

\begin{frame}
	\frametitle{Control variables}
	\begin{itemize}
		\item M\&As: Learning spillover that is separate from the disruption caused.
		\item (4 out of 6 models) Year effect.
		\item (2 out of 6 models) Consolidation and extension as two separate variables, measured more precisely than IV.
		\note {Controlling for consolidation and extension to ensure that we are really capturing the effect of strategic adjustment, rather than effect of the specific direction.\medskip}
		\item Controlling for age and extent of pipeline network.
		\note {Average age, or miles by decade constructed.}
	\end{itemize}
\end{frame}